---
title: "RMDL Assessment 2: Data Analysis Report"
authors:
  - name: B281160
bibliography: b281160ref.bib
execute: 
  echo: true
---

## 1. Background

The Stroop test has become one of the most popular psychological tests to evaluate language control and the scope of the so-called bilingual advantage in the developmental linguistics literature. In summary (see MacLeod & MacDonald 2000 for an effective review) the test captures the subtle processing delay that occurs when we process competing and conflicting stimuli, most frequently realised as a word-colour test where participants read a word and respond with the font colour. Specifically, the Stroop test allows researchers to assess subtle differences in inhibition control and executive functions during automatic tasks, such as reading, when assessing the extent of bilingual lexical access. Despite being used extensively in bilingual acquisition research for decades, the Stroop task remains relevant as ever, especially given substantial new research findings relating to language non-selectivity (Kroll, 2012; Blumenfeld & Marian, 2007). One well-established model for language non-selectivity in bilinguals and multilinguals is The Bilingual Interactive Activation model (BIA+, Dijkstra & Van Heuven 2002) which posits that a bilingual individual’s L1 and L2 is simultaneously activated due to an integrated L1 and L2 lexicon. Bilingual speakers must therefore suppress non-target languages (Kroll, 2008) to prevent cross linguistic interference from this shared lexicon. Consequently, the Stroop task remains an effective method of testing inhibitory control, automaticity, and cross-linguistic interference as a means of understanding non-selective activation in greater depth. This, rather than the notion of bilingual advantage, will be the broad focus of this study.

Much of this research has focussed on the role of age of acquisition (Bugg, 2007), proficiency (Mägiste, 1985) and direction of language switching (Liu et al., 2019) on a bilingual’s ability to suppress information from the non-target language. Despite remaining a contentious issue, the role of orthographic similarity is discussed far less frequently in the literature. According to the BIA+, orthographic similarity between languages may modulate the degree of cross linguistic activation by determining the amount of cross-linguistic bottom-up activation (Dijkstra & Van Heuven, 2002; Van Heuren & Coderre, 2014). Specifically, logographic languages appear to activate language centres in the brain bilaterally where alphabetic languages more predominantly activate left-hemisphere language systems, and thus cross-script language studies provide an opportunity to better understand the neural representations of lexical processing (Van Heuren & Coderre, 2014) and how script type modulates non-selective language activation.

However, previous Stroop experiments investigating the impact of cross-script language interference yield varying results. While some research indicates that script differences may provide stronger cues to suppress cross-linguistic interference, resulting in a smaller Stroop effect for different-script bilinguals (Van Heuven & Coderre, 2014; Bosworth et al., 2021) other studies maintain that orthographic differences have no systematic relationship with the size of the Stroop effect (Lee et al., 1992) and dispute the validity of Van Heuven & Coderre (2014)’s findings (Paap et al., 2015). Paap et al. (2015) note that the study – and the debate as a whole - suffers from no systematic way of measuring orthographical and phonological language similarity, nor the complex ways these two factors interact. Although designing a way to objectively measure these qualities is beyond the scope of the current paper, it highlights that more work is needed to address the ambiguous conclusions regarding orthographic similarity in cross-linguistic interference.

Disentangling the effects of orthographical similarity and phonological similarity is complex, and few natural languages provide the opportunity to explore the effect of orthographic similarity alone. One possible solution is to test bilingual speakers of languages which use more than one script, and explore whether there is an interference effect based on the script alone. Hindi-English bilinguals provide a unique opportunity to explore this problem; it is also relatively understudied in the existing literature, particularly with regards to the effect of orthography on non-selective activation. Hindi is typically written in the alpha-syllabary Devanagari script and shares no phonological system or cognates with English (Mishra & Singh, 2014). However, it is increasingly common in the digital age for Hindi speakers to read and write online using Latin script (Wolf-Sonkin et al., 2019) for convenience. In this scenario, the Latin Hindi word is phonologically and lexically identical, but the use of Latin script creates an orthological overlap with a Latin alphabet language, such as English. In Hindi-English bilinguals, this potentially poses an issue where the retrieval cue determined by the script, as proposed by the BIA+ account (Dijkstra & Van Heuven 2002), is no longer a reliable lexical cue. It is therefore possible that Hindi-English bilinguals experience a greater Stroop effect in Latin script Hindi.

NB: please note that this background draws from work in B281160 RMDL Assignment 1.

### 1.1 Research Questions/Hypotheses

To this end, Hindi provides a unique opportunity to assess the effect of orthographic differences in inhibition control and lexical access in a within-language design. Although there are a small number of tests assessing the Stroop effect in Hindi-English bilinguals (Datta et al., 2019) there is no precedent in the previous literature assessing the Stroop effect in different scripts of the same language. This experiment therefore presents a novel opportunity to explore how inhibitory control is modulated by orthography in an underrepresented area of the previous research, and to understand the effect of orthography in cross-script bilinguals where the phonology is held constant.

Subsequently, the present study asks the following research questions:

1.  Do Hindi-English bilinguals exhibit a stronger Stroop interference effect (i.e., the form of longer reaction times) when reading Latin script Hindi words versus reading Devanagari Hindi words?
2.  What is the difference in the magnitude of interference (in the form of reaction times) between Devanagari Hindi and English conditions, Latin Hindi and English conditions, and Devanagari Hindi and Latin Hindi conditions?

In line with the BIA+ model, we hypothesise that the greatest Stroop effect will be found in the Latin script Hindi conditions as the script will no longer be an effective retrieval cue for non-target language suppression. Subsequently, this would also suggest that the Latin Hindi vs Devanagari Hindi conditions, where the phonology between trials is identical but the script is different, will show the largest degree of interference between all 3 condition comparisons as the phonological representations will compete with the orthography.

## 2. Methodology

### 2.1 Participants

The study sample consists of 5 participants between 18 and 40 years of age; please see Table 1 for an overview of participant characteristics. The sample group was smaller than initially planned due to unexpected constraints with participant recruitment. The participants are highly proficient sequential bilingual speakers of Hindi and English. All participants acquired Hindi at home before the age of 5 and acquired English at school from the age of 5 onwards. Note, for the purposes of this paper that bilingual in this paper is defined as “a person who knows more than one language and uses them to communicate is considered a bilingual, even if she does not have native-like proficiency in both languages” (pp.2 Sorace et al., 2023). The participants did not disclose any clinical history of colour blindness, head trauma, neurological disease, substance abuse, or sedative use in the 72 hours prior to taking the experiment. This was to eliminate any neurological conditions that would interfere with the participants’ reaction times.

### 2.2 Materials

Participants completed an online version of the traditional colour-word Stroop task in Hindi, English, and Latin script Hindi conducted through Testable (Rezlescu, 2020). The stimuli words consisted of red, yellow, blue and green in each language condition, as seen in Table 2 below. Participants responded using the up, down, left, and right keys on the keyboard. In alignment with Datta et al. (2019) only four colours were chosen.

| Hindi | English | Latin script Hindi |
|-------|:--------|:-------------------|
| लाल   | Red     | Laal               |
| पील   | Yellow  | Peela              |
| हरा   | Green   | Hara               |
| नील   | Blue    | Neela              |

: *Table 2: Stroop stimuli by language condition*

| Hindi congruent | Hindi incongruent | English congruent | English incongruent | Latin Script Hindi Congruent | Latin Script Hindi Incongruent |
|------------|------------|------------|------------|------------|------------|
| [लाल]{style="color: red;"} | [लाल]{style="color: blue;"} | [RED]{style="color:red;"} | [RED]{style="color:#F4BB29;"} | [LAAL]{style="color:red;"} | [LAAL]{style="color: green;"} |
| [पीला]{style="color:#f4bb29"} | [पीला]{style="color:green"} | [YELLOW]{style="color:#f4bb29"} | [YELLOW]{style="color:red"} | [PEELA]{style="color:#f4bb29"} | [PEELA]{style="color:blue"} |
| [हरा]{style="color:green"} | [हरा]{style="color:red"} | [GREEN]{style="color:green"} | [GREEN]{style="color:blue"} | [HARA]{style="color:green"} | [HARA]{style="color:#f4bb29"} |
| [नीला]{style="color:blue"} | [नीला]{style="color:#f4bb29"} | [BLUE]{style="color:blue"} | [BLUE]{style="color:green"} | [NEELA]{style="color:blue"} | [NEELA]{style="color:red"} |

: *Table 3: Congruent and incongruent materials for each language condition*

### 2.3 Procedure

Prior to beginning the experimental tasks, participants first completed a consent form informing them of any associated risks or ethical considerations. This was followed by the LEAP-Q questionnaire to confirm their proficiency and age of acquisition of both languages. Participants were emailed a link to the experiment on Testable, with a randomised participant code for anonymity, and invited to complete the test at their convenience in a quiet space where they can concentrate uninterrupted for the duration of the test.

The experiment began with instructions in English explaining the basic premise of the Stroop task: colour words in English, Hindi, and Latin script Hindi will appear in the middle of the screen and participants should respond with the colour of the font, not the colour represented by the word, as quickly as possible using LEFT, RIGHT, UP and DOWN on their keyboard to represent RED, YELLOW, GREEN and BLUE respectively. To familiarise themselves with the process, they completed a short training phase of 10 items where they respond with the font colour of a non-lexical item “XXXXX” and receive written feedback confirming their accuracy. The accuracy and reaction time scores for the training phase were not assessed. They then progressed to the experimental Stroop tasks.

These were divided into three main experimental blocks. The first experimental block assesses the Stroop effect in English, followed by the second in Hindi, and finally the third in in Latin script Hindi. In each block, there were 12 congruent items, where the font colour matches the word, and 12 incongruent items, where the font colour does not match the word. Please see Table 2 for an example of congruent versus incongruent conditions. Following the instructions in the training phase, participants were instructed to respond on their keyboard with the colour of the font and not the colour represented by the word on screen, with LEFT for red, RIGHT for yellow, UP for green and DOWN for blue. The 72 experimental items in each language condition will be pseudo-randomised to ensure the same colours and words do not follow consecutively. There will be a 5-minute break between each experiment. The Stroop task lasted no more than 5 minutes. The key outcome of interest for this study are reaction times, recorded in milliseconds (ms) for each participant and each item across all three experimental blocks. Accuracy is a secondary variable of interest for the English and Latin script Hindi conditions.

### 2.4 Statistical analysis

A Bayesian Lognormal regression analysis will be run in R using the brms program (v4.1.2; R Core Team 2021) to assess the relationship between reaction times, language condition, and congruent or incongruent conditions. Incongruent conditions eliciting the Stroop effect are of particular interest, but the analysis includes congruent conditions to explore the effect of orthography across language conditions more globally.

This is an entirely within-participant experimental design to mitigate for between group effects, therefore it is essential to fit a varying effects model which will control for individual variation in each experimental trial.

RT \~ language + Trialtype + (language \* Trialtype \| participant_ID)

This structure models reaction times (RT) as a dependent variable of trial language and congruency. As all participants were exposed to the same trials, language and trial type were included as varying slopes.

Regarding the research questions, the statistical analysis will be used as follows:

*1. "Do Hindi-English bilinguals exhibit a stronger Stroop interference effect (i.e., the form of longer reaction times) when reading Latin script Hindi words versus reading Devanagari Hindi words?"*

Reaction times in the **incongruent** Latin script Hindi conditions and the Devanagari Hindi will first be compared to determine whether the use of Latin script creates orthological interference when phonology and lexical content of the words are identical. Key summary measures (i.e., mean, median, mode, sd) of the raw data and the regression data will be calculated to compare these.

*"2. What is the difference in the magnitude of interference (in the form of reaction times) between Devanagari Hindi and English conditions, Latin Hindi and English conditions, and Devanagari Hindi and Latin Hindi conditions?"*

Summary measures of reaction times (as described above) will be compared between congruent and incongruent trials across all language conditions to establish the largest within-language Stroop effect. We will then calculate the probability of difference between incongruent trials in Devanagari and Latin Hindi, Latin Hindi and English, and English and Devanagari Hindi to establish the magnitude of interference.

### 2.5 Open Research Statement

This project is fully committed to Open Research practices.

All project items and relevant data are available through the below GitHub repository: https://github.com/sa-lawrence/RMDL2_B281160

Anonymised participant data from Testable can be found under /results/. A .csv version of the Testable experiment is available under /materials/ along with the LEAP-Q questionnaire for language experience and proficiency.

## 3. Results

*Remember to include all code and code output in the submission.*

### 3.1 Workspace set-up

```{r}
#| label: package-load

library(tidyverse)
library(brms)
library(tidybayes)
library(ggdist)
library(posterior)
library(bayesplot)
library(tinytex)


```

### 3.2 Data preparation

First, the Testable files will be read and combined into a single tibble for processing. Note that the first 3 rows of each raw data file were eliminated to remove Testable instructions for cleaner analysis.

```{r}
#| label: read-data
files_to_read = list.files(
  path = "./Data/",
  pattern = "*.csv$",
  recursive = TRUE,
  full.names = TRUE
)

data_lst <- lapply(files_to_read, function(x) {
  read_csv(x, skip = 3) |>
    mutate(
      participant_ID = response[1]
    )
})


full_RMDLSTROOP <- bind_rows(data_lst)

```

Next, the data was filtered to remove filler conditions from the testing phase and condition columns were renamed for clarity. All relevant columns were then selected and processed as a new tibble ("Stroop_filter") which will be used to calculate the raw data summary measures.

```{r}
#| label: filter-data
#NB: Trialtype includes congruent and incongruent trials. Congruent condition is where the font colour and colour word are the same, incongruent is where the font colour and colour word are different. Language refers to the language/script condition. The correct column reflects the accuracy of the participants' answers. RT is reaction time in ms. 

Stroop_filter <- full_RMDLSTROOP |>
  filter(condition1 %in% c("congruent", "incongruent"))

Stroop_filter <- Stroop_filter |>
  rename(Trialtype = condition1, language = condition2)

Stroop_filter <- Stroop_filter |>
  select(rowNo, type, stim1, textColor, Trialtype, language, key, response, RT, 
         correct, participant_ID)
```

### 3.3 Raw data summary measures

As stated in Section 2.4, the key summary measures of interest are the mean, median, standard deviation (SD), and minimum/maximum ranges of the participants' reaction times, grouped by trial type and language.

```{r}
#| label: summary-RT
#Calculating summary measures. 1. Mean and SD of reaction times, grouped by trial type and language
Stroop_filter |> 
  group_by(language, Trialtype) |> 
  summarise(
    mean = mean(RT), sd = sd(RT), median(RT), min(RT), max(RT)) |>
  knitr::kable(digits = 2, col.names = c("Trial Language", "Congruency 
                                         Condition", "Mean RT", "SD RT", 
                                         "Median RT", "Min RT", "Max RT"), 
               caption = "Table 4: Summary measures of reaction times (RT) 
               across language and congruency conditions")
```
Surprisingly, mean RTs in the Latin script Hindi condition were the fastest across incongruent and congruent trials. Within the congruent conditions, English had the highest mean RT at 1559ms followed by Devanagari Hindi at 1261.68ms and 1253.12ms for Latin Hindi. Within the incongruent conditions, Devanagari Hindi was the slowest at 1820.48ms, followed by English at 1811.48ms and Latin Hindi at 1201.97ms. This means the mean RTs for Latin script Hindi were actually faster in the incongruent trials than the congruent trials. Latin Hindi also had the lowest median RTs (congruent = 941ms, incongruent = 964ms) of all language conditions and the lowest SD (congruent = 852.46, incongruent = 624.77), the latter of which indicates less variance than the other two language conditions.



```{r}
#| label: summary-accuracy-pptID
Stroop_filter |>
  group_by(participant_ID, language, Trialtype) |> 
  summarise(
    prop_accuracy = mean(correct)*100) |>
  knitr::kable(digits = 2, col.names = c("Participant ID", "Language", "Congruency", "Accuracy %"), 
               caption = "Table 5: Summary measures of accuracy 
               across language and congruency conditions by participant")
```


```{r}
#| label: summary-accuracy-pptALL
Stroop_filter |>
  group_by(language, Trialtype) |> 
  summarise(
    mean(correct)*100) |>
  knitr::kable(digits = 2, col.names = c("Language", "Congruency", "Mean accuracy %"), 
               caption = "Table 6: Summary measures of accuracy 
               across language and congruency conditions")
```

Tables 5 and 6 show the accuracy of the participants' responses across language conditions and congruency, either by participant average (Table 5) or grouped by language and congruency condition (Table 6). These show that participants overall were the least accurate in the Devanagari Hindi incongruent condition, and most accurate in the congruent English conditions.

### 3.4 Plots: Raw data summaries

```{r}
#| label: raw-jitter-violin
#Plotting the raw RT data. 

jitterviolinRT_raw <- Stroop_filter |>
  ggplot(aes(language, RT, fill = language)) +
  geom_jitter(alpha = 0.2, width = 0.1) +
  geom_violin(width = 0.7) +
  scale_fill_brewer(palette = "Dark2") +
  facet_grid(cols = vars(Trialtype)) +
  labs(
    title = "Figure 1: Mean RTs by language and congruency",
    x = "Trial language",
    y = "Response Reaction Time(ms)"
  )

jitterviolinRT_raw
```
This is a jitter plot of RTs grouped by language and separated by congruency condition in two panels. Within each panel, the participant’s language is represented by colour, with green for English, orange for Hindi, and purple for Latin Hindi. The jitter plots show the individual raw data points, and the violin overlay indicates the density and variability in participants' responses. As shown in the summary conditions above, there appears to be greater variability in the RTs in  incongruent conditions for English and Devanagari Hindi, with the exception of the outliers in the English congruent condition, reporting RTs of over 8000ms. Latin Hindi is the exception to this pattern, with similar density width patterns yet less overall variability in RT range. 

```{r}
# This jitter plot depicts the mean accuracy percentage for each participant grouped by language and trial type to explore the relationship between accuracy, language, and trial type.
#| label: raw-jitter-accuracy

accuracy_stroop_filter <- Stroop_filter |>
  group_by(participant_ID, Trialtype, language) |> 
  summarise(
    prop_accuracy = round(mean(correct),2)) 

ggplot() +
  geom_jitter(
    data = accuracy_stroop_filter, aes(language, prop_accuracy),
    width = 0.1, alpha = 0.5
  ) + 
  facet_grid(cols = vars(Trialtype)) +
  stat_summary(
    data = Stroop_filter,
    aes(language, correct, colour = language),
     fun.data = "mean_cl_boot", size = 0.5
  ) +
  labs(
    title = "Figure 2: Percentage of mean accuracy in Stroop task by participant",
    caption = "Mean proportion is represented by coloured points with 95% bootstrapped Confidence Intervals.",
    x = "Language",
    y = "Accuracy"
  ) +
  theme(legend.position = "none")
  facet_grid(cols = vars(Trialtype)) 
  
```
Figure 2 is a jitter plot depicting the mean accuracy rating per participant by language and congruency condition. Mean accuracy scores were calculated per participant (as seen in section 3.3). Then, means of these means and confidence intervals were fitted with the stat_summary() function and plotted with geom_jitter(). As with Figure 1 above, congruency and incongruency are separated into 2 panels to assess the difference in Stroop effect from the incongruent conditions. English data is in red, Hindi is in green, and Latin Hindi is in blue. Accuracy between congruent and incongruent varied most in the English condition, while accuracy in Latin Hindi remains consistent between congruency conditions. Incongruent Hindi trials shows the greatest variability in accuracy, with a possible outlier to be considered. 

### Summarise the data

*Calculate and report appropriate summary measures for relevant variables in the data.*

```{r}

#Calculating summary measures. 1. Mean and SD of reaction times, grouped by trial type and language
raw_stroop_filter <- Stroop_filter |> 
  group_by(Trialtype, language) |> 
  summarise(
    mean = mean(RT), sd = sd(RT))

accuracy_stroop_filter <- Stroop_filter |>
  group_by(participant_ID, Trialtype, language) |> 
  summarise(
    prop_accuracy = round(mean(correct),2))


```

## 4. Lognormal regression modelling

As stated in the statistical analysis section (2.4) this data will be analysed using a Bayesian Lognormal regression model using the brms function. This model was chosen for two key reasons: firstly, the lognormal distribution family is most appropriate for variables that are continuous and only positive or above 0, and is therefore well suited to analyse reaction times, and secondly, RTs tend to be right skewed as they are bound to positive numbers, so they need to be logged to approximate a Gaussian distribution.

This model is designed to assess the relationship between RTs, language script, and congruency conditions to see how the Stroop effect and longer reaction times are affected by either independent variable. Crucially, it was fitted with varying slopes to account for inter-participant variability. As this is a within-participant design, the within participant variables to include in the slopes are the same as the main model. Note that, as accuracy and proficiency are not of primary relevance to our research questions, these were not included in the models below.

Finally, the model was fitted with 0.999 adapt_delta and a max_treedepth of 15 to avoid divergent transitions.

```{r}
#running the lognormal regression model, as the variable of interest is continuous and cannot be below 0. If you get divergent transitions, add a new parameter to brm model. 

ENHIstroopbrm <- brm(
  RT ~ language * Trialtype + (language * Trialtype | participant_ID),
  family = lognormal,
  data = Stroop_filter,
  cores = 4,
  seed = 6789,
  file = "data/cache/ENHIstroopbrm8",
  control = list(adapt_delta = 0.999, max_treedepth = 15)
)
```

```{r}
summary(ENHIstroopbrm)
```

```{r}
#draws so we can pull summary measures
stroopbrm_draws <- as_draws_df(ENHIstroopbrm)
stroopbrm_draws

#Mutating the draws to get the intercepts for all six conditions of interest for reaction times: English Congruent, English Incongruent, Hindi Congruent, Hindi Incongruent, Latin Hindi Congruent and Latin Hindi Incongruent.
stroopbrm_draws2 <- stroopbrm_draws |> 
  mutate(
    english_congruent = b_Intercept,
    hindi_congruent = b_Intercept + b_languagehindi,
    latin_hindi_congruent = b_Intercept + b_languagelatinhindi, 
    english_incongruent = b_Intercept + b_Trialtypeincongruent,
    hindi_incongruent = b_Intercept + b_languagehindi + b_Trialtypeincongruent + `b_languagehindi:Trialtypeincongruent`,
    latin_hindi_incongruent = b_Intercept + b_languagelatinhindi + b_Trialtypeincongruent + `b_languagelatinhindi:Trialtypeincongruent`
  )
    
```

```{r}
#As this is a log-normal regression model, I am unlogging the RTs to convert them back to ms.

stroopbrm_draws3 <- stroopbrm_draws2 |>
  select(english_congruent, hindi_congruent, latin_hindi_congruent, english_incongruent, hindi_incongruent, latin_hindi_incongruent) 


stroopdrawsunlog <- stroopbrm_draws3 |>
  mutate(
    english_congruent_log = english_congruent,
    hindi_congruent_log = hindi_congruent,
    latin_hindi_congruent_log = latin_hindi_congruent,
    english_incongruent_log = english_incongruent,
    hindi_incongruent_log = hindi_incongruent,
    latin_hindi_incongruent_log = latin_hindi_incongruent,
    english_congruent_unlog = exp(english_congruent),
    hindi_congruent_unlog = exp(hindi_congruent),
    latin_hindi_congruent_unlog = exp(latin_hindi_congruent),
    english_incongruent_unlog = exp(english_incongruent),
    hindi_incongruent_unlog = exp(hindi_incongruent),
    latin_hindi_incongruent_unlog = exp(latin_hindi_incongruent)
  )


```

```{r}
#pivot longer UNLOGGED VALUES
stroopdrawsunlog_long <- stroopdrawsunlog |> 
  select(english_congruent_unlog, english_incongruent_unlog, 
         hindi_congruent_unlog, hindi_incongruent_unlog, latin_hindi_congruent_unlog, latin_hindi_incongruent_unlog) |> 
  pivot_longer(everything(), names_to = "condition", values_to = "RT_ms")

#summarise pivot longer 
stroopdrawsunlog_long_tab <- stroopdrawsunlog_long |> 
  group_by(condition) |> 
  summarise(
    mean = mean(RT_ms), sd = sd(RT_ms),
    '95%' = paste0("[", paste(quantile(RT_ms, c(0.025, 0.975)) |> round(), collapse = ", "), "]"),
    '66%' = paste0("[", paste(quantile(RT_ms, c(0.17, 0.83)) |> round(), collapse = ", "), "]")
  )

stroopdrawsunlog_long_tab
```

```{r}
conditional_effects(ENHIstroopbrm, "language:Trialtype")
```

## Discussion

*Write a brief discussion of the your model results in light of the research question.*
