---
title: "RMDL Assessment 2: Data Analysis Report"
authors:
  - name: B281160
bibliography: references.bib
execute: 
  echo: true
---

## Background

The Stroop test has become one of the most popular psychological tests to evaluate language control and the scope of the so-called bilingual advantage in the developmental linguistics literature. In summary (see MacLeod & MacDonald 2000 for an effective review) the test captures the subtle processing delay that occurs when we process competing and conflicting stimuli, most frequently realised as a word-colour test where participants read a word and respond with the font colour. Specifically, the Stroop test allows researchers to assess subtle differences in inhibition control and executive functions during automatic tasks, such as reading, when assessing the extent of bilingual lexical access. Despite being used extensively in bilingual acquisition research for decades, the Stroop task remains relevant as ever, especially given substantial new research findings relating to language non-selectivity (Kroll, 2012; Blumenfeld & Marian, 2007). One well-established model for language non-selectivity in bilinguals and multilinguals is The Bilingual Interactive Activation model (BIA+, Dijkstra & Van Heuven 2002) which posits that a bilingual individual’s L1 and L2 is simultaneously activated due to an integrated L1 and L2 lexicon. Bilingual speakers must suppress non-target languages (Kroll, 2008) to prevent cross linguistic interference from this shared lexicon. Consequently, the Stroop task remains an effective method of testing inhibitory control and automaticity as a means of understanding this non-selective activation in greater depth.

Much of this research has focussed on the role of age of acquisition (Bugg, 2007), proficiency (Mägiste, 1985) direction of language switching (Liu et al., 2019) and executive control. However, this has historically focussed on languages from the global north which use the Latin alphabet, resulting in a relatively small minority of studies which investigate the role of cross-script language switching on these mechanisms. Cross-script language studies crucially provide an opportunity to better understand the neural representations of lexical processing, as logographic languages appear to activate language centres in the brain bilaterally where alphabetic languages more predominantly activate left-hemisphere language systems (Van Heuren & Coderrre, 2025). However, previous Stroop experiments investigating the impact of cross-script language interference yield varying results. While some researchers maintain that orthographic differences have no systematic relationship with the size of the Stroop effect (Lee et al., 1992) more recent studies such as that of Van Heuven & Coderre (2025) which assessed different-script trilinguals suggest that script differences may provide stronger cues to support language suppression and avoid cross-linguistic interference, resulting in a smaller Stroop effect. Consequently, more work is needed to resolve the contradictions surrounding the impact of cross-script differences in inhibition control.

One language which is understudied in the existing literature, particularly with regards to Stroop tests, is Hindi. While Hindi is typically written in Devanagari script, it is increasingly common in the digital age for Hindi speakers to read and write online or in instant messages using Latin script (Wolf-Sonkin et al., 2019) for convenience. To this end, Hindi provides a unique opportunity to assess the effect of orthographic differences in inhibition control and lexical access in a within-language design. This presents the research question addressed by the present study: is the Stroop effect in Hindi-English bilinguals modulated by different scripts? As a result, this study proposes to investigate the difference in Stroop effect between traditional Devanagari Hindi, English, and Latin script Hindi among Hindi-English bilinguals to isolate the effect of orthographic differences in bilingual lexical access.

Although there are a small number of tests assessing the Stroop effect in Hindi-English bilinguals (Datta et al., 2019) there is no precedent in the previous literature assessing the Stroop effect in different scripts of the same language. This experiment therefore presents a novel opportunity to explore variation in inhibitory control modulated by script in an underrepresented area of the previous research.

### Research Questions/Hypotheses

## Methods

*Briefly report the methods of the study (participants, materials, procedure, statistical analysis...). Note that when discussing the statistical analysis you should clearly state how the analysis results will be used to assess the RQs/RHs (be as specific and precise as you can). Include an Open Research/Data availability statement.*

## Data

*Remember to include all code and code output in the submission.*

### Read and process the data

*Read the data here. If the data needs processing, like mutating columns or creating new ones, briefly explain why processing is needed and what you did*. *Include a description of each of the relevant columns in the data.*

```{r}
#set up packages

library(tidyverse)
library(brms)
library(tidybayes)
library(ggdist)
library(posterior)
library(bayesplot)

#| label: read-data

# Example code: delete

files_to_read = list.files(
  path = "./Data/",
  pattern = "*.csv$",
  recursive = TRUE,
  full.names = TRUE
)

data_lst <- lapply(files_to_read, function(x) {
  read_csv(x, skip = 3) %>%
    mutate(
      participant_ID = response[1]
    )
})


full_RMDLSTROOP <- bind_rows(data_lst)

```

```{r}
#filtering the data to remove filler conditions, renaming columns for clarity. 

#NB: Trialtype includes congruent and incongruent trials. Congruent condition is where the font colour and colour word are the same, incongruent is where the font colour and colour word are different. Language refers to the language/script condition. The correct column reflects the accuracy of the participants' answers. RT is reaction time in ms. 

Stroop_filter <- full_RMDLSTROOP |>
  filter(condition1 %in% c("congruent", "incongruent"))

Stroop_filter <- Stroop_filter %>%
  rename(Trialtype = condition1, language = condition2)

Stroop_filter <- Stroop_filter %>%
  select(rowNo, type, stim1, textColor, Trialtype, language, key, response, RT, correct, participant_ID)
```

### Plot the data

*Create relevant plots: for each, include a description of the plot (type, axes, colours used and so on) and a description of relevant patterns.*

```{r}
#Plotting the raw data. This is a jitter plot of RTs separated by trial type and script. The two panels indicate whether the trial type was congruent or incongruent. Within each panel, the participant’s language is represented by colour. Congruent conditions had faster RTs than incongruent conditions, and x language was faster than Y language. From the plot it can also be seen that there is greater variability in xxxxx . Within each group of xxxx there does not seem to be any specific relation between RTs and script.

Stroop_filter |>
  ggplot(aes(language, RT, fill = language)) +
  geom_jitter(alpha = 0.2, width = 0.1) +
  geom_violin(width = 0.7) +
  scale_fill_brewer(palette = "Dark2") +
  facet_grid(cols = vars(Trialtype)) +
  xlab("Trial language") +
  ylab("Response Reaction Time(ms)")


```

```{r}
# This jitter plot depicts the mean accuracy percentage for each participant grouped by language and trial type to explore the relationship between accuracy, language, and trial type.
accuracy_stroop_filter <- Stroop_filter |>
  group_by(participant_ID, Trialtype, language) |> 
  summarise(
    prop_accuracy = round(mean(correct),2)) 

ggplot() +
  geom_jitter(
    data = accuracy_stroop_filter, aes(language, prop_accuracy),
    width = 0.1, alpha = 0.5
  ) + 
  facet_grid(cols = vars(Trialtype)) +
  stat_summary(
    data = Stroop_filter,
    aes(language, correct, colour = language),
     fun.data = "mean_cl_boot", size = 0.5
  ) +
  labs(
    title = "Percentage of accuracy in Stroop task",
    caption = "Mean proportion is represented by coloured points with 95% bootstrapped Confidence Intervals.",
    x = "Language",
    y = "Accuracy"
  ) +
  theme(legend.position = "none")
  facet_grid(cols = vars(Trialtype)) 
  
```

### Summarise the data

*Calculate and report appropriate summary measures for relevant variables in the data.*

```{r}
#Calculating summary measures. 1. Mean and SD of reaction times, grouped by trial type and language
raw_stroop_filter <- Stroop_filter |> 
  group_by(Trialtype, language) |> 
  summarise(
    mean = mean(RT), sd = sd(RT))

accuracy_stroop_filter <- Stroop_filter |>
  group_by(participant_ID, Trialtype, language) |> 
  summarise(
    prop_accuracy = round(mean(correct),2))


```

## Regression modelling

*Write an explanation of the regression model(s) you will run to answer the chosen research question. Fit the model and get a model summary and relevant model plots. Report the results of the model.*

```{r}
#running the lognormal regression model, as the variable of interest is continuous and cannot be below 0. If you get divergent transitions, add a new parameter to brm model. 

ENHIstroopbrm <- brm(
  RT ~ language * Trialtype + (language * Trialtype | participant_ID),
  family = lognormal,
  data = Stroop_filter,
  cores = 4,
  seed = 6789,
  file = "data/cache/ENHIstroopbrm7",
  control = list(adapt_delta = 0.999, max_treedepth = 15)
)
```

```{r}
summary(ENHIstroopbrm)
```

```{r}
#draws so we can pull summary measures
stroopbrm_draws <- as_draws_df(ENHIstroopbrm)
stroopbrm_draws

#Mutating the draws to get the intercepts for all six conditions of interest for reaction times: English Congruent, English Incongruent, Hindi Congruent, Hindi Incongruent, Latin Hindi Congruent and Latin Hindi Incongruent.
stroopbrm_draws2 <- stroopbrm_draws |> 
  mutate(
    english_congruent = b_Intercept,
    hindi_congruent = b_Intercept + b_languagehindi,
    latin_hindi_congruent = b_Intercept + b_languagelatinhindi, 
    english_incongruent = b_Intercept + b_Trialtypeincongruent,
    hindi_incongruent = b_Intercept + b_languagehindi + b_Trialtypeincongruent + `b_languagehindi:Trialtypeincongruent`,
    latin_hindi_incongruent = b_Intercept + b_languagelatinhindi + b_Trialtypeincongruent + `b_languagelatinhindi:Trialtypeincongruent`
  )
    
```

```{r}
#As this is a log-normal regression model, I am unlogging the RTs to convert them back to ms.

stroopbrm_draws3 <- stroopbrm_draws2 |>
  select(english_congruent, hindi_congruent, latin_hindi_congruent, english_incongruent, hindi_incongruent, latin_hindi_incongruent) 


stroopdrawsunlog <- stroopbrm_draws3 |>
  mutate(
    english_congruent_log = english_congruent,
    hindi_congruent_log = hindi_congruent,
    latin_hindi_congruent_log = latin_hindi_congruent,
    english_incongruent_log = english_incongruent,
    hindi_incongruent_log = hindi_incongruent,
    latin_hindi_incongruent_log = latin_hindi_incongruent,
    english_congruent_unlog = exp(english_congruent),
    hindi_congruent_unlog = exp(hindi_congruent),
    latin_hindi_congruent_unlog = exp(latin_hindi_congruent),
    english_incongruent_unlog = exp(english_incongruent),
    hindi_incongruent_unlog = exp(hindi_incongruent),
    latin_hindi_incongruent_unlog = exp(latin_hindi_incongruent)
  )


```

```{r}
#pivot longer UNLOGGED VALUES
stroopdrawsunlog_long <- stroopdrawsunlog |> 
  select(english_congruent_unlog, english_incongruent_unlog, 
         hindi_congruent_unlog, hindi_incongruent_unlog, latin_hindi_congruent_unlog, latin_hindi_incongruent_unlog) |> 
  pivot_longer(everything(), names_to = "condition", values_to = "RT_ms")

#summarise pivot longer 
stroopdrawsunlog_long_tab <- stroopdrawsunlog_long |> 
  group_by(condition) |> 
  summarise(
    mean = mean(RT_ms), sd = sd(RT_ms),
    '95%' = paste0("[", paste(quantile(RT_ms, c(0.025, 0.975)) |> round(), collapse = ", "), "]"),
    '66%' = paste0("[", paste(quantile(RT_ms, c(0.17, 0.83)) |> round(), collapse = ", "), "]")
  )

stroopdrawsunlog_long_tab
```

```{r}
conditional_effects(ENHIstroopbrm, "language:Trialtype")
```

## Discussion

*Write a brief discussion of the your model results in light of the research question.*
